---
title: "Outcome of Dogs and Cats Admitted to an Animal Shelter in a No Kill city"
author: "Leigh Tanji, Carter Kyle, Naila Arsky"
date: "2018/03/08"
output: html_document
---
*git repository used: https://github.com/ltanji21/SDS201-Project/blob/master/SDS201%20Project.Rmd *


### Purpose

The purpose of our project is to determine the correlations between sex, color, species, breed, age and fertility and the adoption of cats and dogs from Animal Center, a domestic shelter in Austin Texas, a no kill city.

The general topic we want to study is the relationship between specific physical characteristics of animals and the rate of adoption. We plan to investigate whether or not some of these characteristics are correlated and whether or not this implies causation between these variables. If there is correlation between variables, we will investigate confounding variables. Our research may shed light on our hypothesis that animals with lighter colors tend to have a higher adoption rate, and that certain physical characteristics in a breed also affect the probability of an animal finding a home. Although we believe all of these factors are true according to common knowledge, the outcomes of this study may or may not confirm these beliefs.


### Data

The data we are using is called Austin Animal Center Outcomes. It can be found at https://catalog.data.gov/dataset/austin-animal-center-outcomes-version-1-demo. This data set contains the outcome for animals taken in by animal shelters in Austin, Texas as well as the amount of time the animal was held in the shelter and physical descriptors of the animals such as age and breed. For this analysis, we will only observe animals that were under the `Outcome.Type = "Adopted"` and also use Logistic Regression.

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(oilabs)
library(mosaic)
library(tidyr)

Austin_Animals <- read.csv("https://data.austintexas.gov/api/views/9t4d-g238/rows.csv?accessType=DOWNLOAD")
```

## Data Dictionary

```{r}
#Cleaning Up data.
Austin_Animals2 <- Austin_Animals %>%
  filter(!Outcome.Type == "Rto-Adopt",
         !Outcome.Type == "Return to Owner",
         Sex.upon.Outcome!= "NULL",
         Sex.upon.Outcome!= "Unknown",
         Animal.Type %in% c("Cat", "Dog")) %>%
  mutate(Homed = ifelse(Outcome.Type == "Adoption", 1, 0)) %>%
  select(DateTime, MonthYear, Date.of.Birth, Homed, Outcome.Subtype, Animal.Type, Sex.upon.Outcome, Age.upon.Outcome, Breed, Outcome.Type, Color) %>%
  mutate(Black = ifelse(Color %in% c("Black", 
                                   "Black/Black", 
                                   "Black Brindle/Black",
                                   "Black Smoke",
                                   "BLack Smoke/Black",
                                   "Black Tabby/Black",
                                   "Black/Black Brindle",
                                   "Black/Black Smoke",
                                   "Black/Black Tabby"), 1, 0)) %>%
  separate(Age.upon.Outcome, c("num", "unit"), sep = " " , remove = FALSE) %>%
  mutate(num = as.numeric(num),
         age = ifelse(unit %in% c("day", "days"), num/30, 
                      ifelse(unit %in% c("month", "months"), num, 
                                         ifelse(unit %in% c("week", "weeks"), num/4, 
                                                ifelse(unit %in% c("year", "years"), num*12, NA)))))
```

```{r}
Austin_Animals2 <- Austin_Animals2 %>%
  mutate(Sex = ifelse(Sex.upon.Outcome %>% c("Intact Female", "Spayed Female"), 1, 0)) %>%
  mutate(Fixed = ifelse(Sex.upon.Outcome %>% c("Neutered Male", "Spayed Female"), 1, 0)) %>% 
```


**Str**
The Str describes each of the twelve variables that we have in the dataframe “Adopted”. 

* `Animal.ID` has 73442 levels which correspond to an adopted animal. The variable with the most levels. Interestingly, it has more entries than the dimension of the data frame itself. We are unsure about why this may be since each animal should only have one unique identifier. 

* Name has only 14882 levels since not all animals have names in the shelter.

* DateTime is made up of 67009 levels since some animals may have been taken in by the shelter at the same times.

* Date.of.Birth also only has 5987 levels since some animals may have been born on the same day and year. 

* Outcome.Type has 10 levels (In reality, only has 1 since filtered by “Adoption”)

* Outcome.Subtype is made up of 20 levels corresponding to Barn, Foster, Offsite, etc. Again, the number of outcome subtypes are more than the number of levels. Should only be four. 

* Sex.Upon Outcome is a factor with 6 levels. Actually made up of 5 types: Intact Female, Intact Male, Neutered Male, Spayed Female, and Unknown. 

* Age.Upon.Outcome is a Factor with 47 levels. Not sure why there are so many levels since it’s not categorized.

* Breed has 2183 levels. In reality made up of 1543 breed lists.

* Color has 533 levels, again has 2768 colors of animals. 

** Amount of the levels don’t seem to match the actual numbers in the variables. 
```{r}
str(Austin_Animals2)
```

* *Favstats*
The mean age upon outcome for adopted animals was 25 months. The oldest animal adopted was 45 months old and the youngest was 1 month old. 

* *Outcome.Subtype*
Most of the adopted animals did not have a subtype: 82%. The variable subtype had 4 levels.

* *Animal.Type*
The animal with the highest rate of adoption was dog, at 64% of all adopted animals. This variable had 5 levels.

* *Sex.Upon Outcome*
Neutered males had the highest adoption rate, at 48% of all adopted animals. This variable had 5 levels. There were 70 animals whose sex was not known upon adoption. 

* *Breed*
There are 1,543 levels for this variable. However, because all shelter animals are stray, the breeds of the parents of each animal are unknown. Therefore, we cannot be certain that the variable “Breed” is always accurate.

* *Color*
There are 393 levels for this variable.

```{r}
#The one quantitative variable
favstats(~Age.upon.Outcome, data = Austin_Animals2)

# six other categotical variables
Austin_Animals2 %>%
  group_by(Outcome.Subtype) %>%
  tally()

Austin_Animals2 %>%
  group_by(Animal.Type) %>%
  tally()

Austin_Animals2 %>%
  group_by(Sex.upon.Outcome) %>%
  tally()

Austin_Animals2 %>%
  group_by(Breed) %>%
  tally()

Austin_Animals2 %>%
  group_by(Color) %>%
  tally()

Austin_Animals2 %>%
  group_by(Homed) %>%
  tally()
```


### Population

The observational units are dogs and cats that were taken into animal shelters in Austin, Texas. The larger population we are generalizing to is all dogs and cats in animal shelters in Austin, Texas. It has 79.4K rows and 12 columns. Each row represents an animal and the columns include animal ID, name, date time, monthyear, date of birth, outcome, outcome subtype, breed, animal type, sex, age upon outcome, and color. 

### Response Variable
Adoption rate in percentage. We believe that 80% of spayed/neutered animals, 50% female/male, and 50% animal types (i.e. cat vs. dog), were adopted. Within breeds, we believe that pitbulls and Japanese chins for dogs and sphinxes for cats will have lower percentages of adoption. Black coloured animals are also considered as having lower adoption rates. 

### Explanatory Variables

Our explanatory variables are sex, color, animal type, breed, whether or not they were sterilized(spayed and neutered), and the age of the animal. The possible categories for sex are male and female. A few of the categories for dog breed are Labrador, Chihuahua, and Akita . A few of the categories for cat breed are domestic shorthair, domestic mediumhair, and ragdoll mix. A few of the possible categories for cat color are tabby, calico, and black. A few of the possible categories for dog color are cream, red, and black. The possible categories for animal type are dogs and cats. The possible categories for sterilized are sterilized and not sterilized. The units for age are months. All explanatory variables were made using physical observations of the animals 

### Analysis

```{r}
#sex, color, species, breed, age and fertility

```



















