---
title: "Outcome of Dogs and Cats Admitted to an Animal Shelter in a No Kill city"
author: "Leigh Tanji, Carter Kyle, Naila Arsky"
date: "2018/03/08"
output: html_document
---
*git repository used: https://github.com/ltanji21/SDS201-Project/blob/master/SDS201%20Project.Rmd *


### Purpose

The purpose of our project is to determine the correlations between sex, color, species, breed, age and fertility and the adoption of cats and dogs from Animal Center, a domestic shelter in Austin Texas, a no kill city.

The general topic we want to study is the relationship between specific physical characteristics of animals and the rate of adoption. We plan to investigate whether or not some of these characteristics are correlated and whether or not this implies causation between these variables. If there is correlation between variables, we will investigate confounding variables. Our research may shed light on our hypothesis that animals with lighter colors tend to have a higher adoption rate, and that certain physical characteristics in a breed also affect the probability of an animal finding a home. Although we believe all of these factors are true according to common knowledge, the outcomes of this study may or may not confirm these beliefs.


### Data

The data we are using is called Austin Animal Center Outcomes. It can be found at https://catalog.data.gov/dataset/austin-animal-center-outcomes-version-1-demo. This data set contains the outcome for animals taken in by animal shelters in Austin, Texas as well as the amount of time the animal was held in the shelter and physical descriptors of the animals such as age and breed. For this analysis, we will only observe animals that were under the `Outcome.Type = "Adopted"`. 

```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(oilabs)
library(mosaic)

Austin_Animals <- read.csv("https://data.austintexas.gov/api/views/9t4d-g238/rows.csv?accessType=DOWNLOAD")

Adopted <- Austin_Animals %>%
  filter(Outcome.Type == "Adopted")
```

## Data Dictionary


```{r}
str(Adopted)
```

```{r}
#The one quantitative variable
favstats(~Age.upon.Outcome, data = Adopted)

# six other categotical variables
Adopted %>%
  group_by(Outcome.Subtype) %>%
  tally()

Adopted %>%
  group_by(Animal.Type) %>%
  tally()

Adopted %>%
  group_by(Sex.upon.Outcome) %>%
  tally()

Adopted %>%
  group_by(Breed) %>%
  tally()

Adopted %>%
  group_by(Color) %>%
  tally()
```


### Population

The observational units are dogs and cats that were taken into animal shelters in Austin, Texas. The larger population we are generalizing to is all dogs and cats in animal shelters in Austin, Texas. It has 79.4K rows and 12 columns. Each row represents an animal and the columns include animal ID, name, date time, monthyear, date of birth, outcome, outcome subtype, breed, animal type, sex, age upon outcome, and color. 

### Response Variable
Adoption rate in percentage. We believe that 80% of spayed/neutered animals, 50% female/male, and 50% animal types (i.e. cat vs. dog), were adopted. Within breeds, we believe that pitbulls and Japanese chins for dogs and sphinxes for cats will have lower percentages of adoption. Black coloured animals are also considered as having lower adoption rates. 

### Explanatory Variables

Our explanatory variables are sex, color, animal type, breed, whether or not they were sterilized(spayed and neutered), and the age of the animal. The possible categories for sex are male and female. A few of the categories for dog breed are Labrador, Chihuahua, and Akita . A few of the categories for cat breed are domestic shorthair, domestic mediumhair, and ragdoll mix. A few of the possible categories for cat color are tabby, calico, and black. A few of the possible categories for dog color are cream, red, and black. The possible categories for animal type are dogs and cats. The possible categories for sterilized are sterilized and not sterilized. The units for age are months. All explanatory variables were made using physical observations of the animals 






















